---
title: "Assignment 1 ST464/ST684"
author: "madhusudan 20250509"
date: "`r format(Sys.time(), '%X %d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=4)
```


```{r, eval=T, echo=FALSE}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(GGally))

```

Question 1
(a)
```{r}
x<-matrix(c(-3,5,-4,0,4,-1,-2,3,2,-5),ncol=2)
x
eud <- as.matrix(dist(x, method = "euclidean"))
eud2 <- eud*eud
eud2
```





#### Question 4



```{r eval=T} 
# change to eval=T
# place the downloaded csv file in the same folder as this Rmd file.
w <- read.csv("C:/Users/Madhusudan/Desktop/Learning/Maynooth University/Semester2/ST684/Assignment/worldhappiness2019.csv")
names(w)<- c( "Rank", "Country", "Score", "GDP", "Family", "Life", "Freedom", "Generosity", "Trust")
rownames(w)<- w$Country
wscores <- w[,-(1:3)]
```

(a)
```{r}
cor(wscores)

#we see that life and GDP is having highest correlation (0.8354)
library(ggplot2)
ggplot(data=wscores, aes(x=GDP, y=Life))+geom_point() +
      labs(title = "Scatterplot of GDP vs Life")
head(wscores)

print(wscores[wscores$Life==0,])
```
The pair of variables with the highest correlation are GDP and Life.
The outlying country on this graph is Swaziland.


(b)
```{r}

#Euclidean matrix of wscores - no standardistaion
e <- dist(wscores,method="euclidean")
head(e)


##countries clustering (Average linkage)
clust <- hclust(e,"average")
clust

#Dendrogram of clustering
d <- as.dendrogram(clust)
plot(d)
```

There is ONE outlier country called Dominican Republican which is on extreme left of cluster dendrogram.


(c)
```{r}
#dividing into 3 clusters 
c <- cutree(clust, 3)
table(c)

#Checking which clusters are having highest score
source('C:/Users/Madhusudan/Desktop/Learning/Maynooth University/Semester2/ST684/Assignment/sumPartition.R')
ws <- sumPartition(wscores,c)

#looking at country from cluster-wise
wsnew <- wscores
wsnew$cluster <- factor(c)
wsnew$country <- rownames(wsnew)
head(wsnew)

#Ireland
wsire <- wsnew[wsnew$country=="Ireland",]
wsire

#India
wsind <- wsnew[wsnew$country=="India",]
wsind
```

The cluster sizes are 111, 44 and 1 for cluster 1,2 and 3 respectively.

Ireland belongs to cluster 1.

India belongs to cluster 2.

Cluster 1 has the highest Family score.

Cluster 1 has the highest  scores overall.


(d)
```{r}
library(GGally)
head(wsnew)
ggparcoord(wsnew, columns = 1:6, groupColumn = "cluster", scale= "globalminmax")

```

The lowest scoring cluster is unusual because it is the cluster having only one country i.e. Dominican Republic. The most probable reason for this is the Family score of DR which is totally different from all other countries.

(e)
```{r}
set.seed(123)
km <- kmeans(wscores, 3, nstart = 10)
km$centers
table(km$cluster)
```
AS we see that there are three groups and there are plentry of countries in each group so
There are no outlier countries


(f)
```{r}
ws_km <- wscores
#creating new column as per factor of kmeans
ws_km$cluster <- factor(km$cluster)
#check new df once
head(ws_km)
#creating plot
ggparcoord(ws_km, columns = 1:6, groupColumn = "cluster", scale= "globalminmax")
```

Cluster 1 (red) has the highest Family score.
Cluster 1 (red) has the highest  scores overall.



{g}
```{r}
#g
ws_km$score <- w$Score
ggplot(ws_km, aes(x=cluster, y=score, color=ws_km$cluster)) + geom_boxplot() +
  labs(title = "Box plot of score for three clusters(color-wise)")

```

We can clearly see that cluster 1 countries are having best scores i.e. they are the happiest while cluster 2 is in middle and cluster 3 countries(least score) are the lease happy countries.
